---
output: rmarkdown::github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      collapse=TRUE, 
                      comment="##", 
                      fig.retina=2, 
                      fig.path = "README_figs/README-")

library(HypeRIMU)
```

[![Build Status](https://travis-ci.org/JohannesFriedrich/HypeRIMU.svg?branch=master)](https://travis-ci.org/JohannesFriedrich/HypeRIMU)
[![Build status](https://ci.appveyor.com/api/projects/status/lm7jn3t558yxveve?svg=true)](https://ci.appveyor.com/project/JohannesFriedrich/hyperimu)
[![Coverage Status](https://codecov.io/gh/JohannesFriedrich/HypeRIMU/branch/master/graph/badge.svg)](https://codecov.io/gh/JohannesFriedrich/HypeRIMU)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)


HypeRIMU is a **R**-package for reading sensor data from the Android app [HyperIMU](https://play.google.com/store/apps/details?id=com.ianovir.hyper_imu). 

# Installation

```{r, eval = FALSE}
if(!require("devtools"))
  install.packages("devtools")
devtools::install_github("JohannesFriedrich/HypeRIMU")
```

# Motivation

Modern smartphones include a lot of sensors, e.g. for gravity or linear acceleration. Some apps provide the data collected from these sensors. One famous android app is "HyperIMU" (IMU: Inertial measurement unit). There are other apps available, but HyperIMU has some advantages:

* Choose which sensor data should be collected
* Choose different streaming protocols: SD-card (locally on the smarphone), UDP, or TCP
* The sampling rate can be adjusted

The **R**-package HypeRIMU provides some handy functions to visualise sensor data. Different streaming protocolls are supported (a local .csv file and TCP).

# Usage

For using the smartphone app, see [HIMUServer](https://github.com/ianovir/HIMUServer/) and [HyperIMU Android](https://play.google.com/store/apps/details?id=com.ianovir.hyper_imu&hl=de) and install the app on your smartphone.

### Read sensor data

Until now two different methods to read the sensor data are supported: TCP and a local file.

#### TCP

The TCP protocol is also supported. Using this method needs to know your IP address to sumbit in the smartphone app, see [HIMUServer](https://github.com/ianovir/HIMUServer/).

```{r, eval=FALSE}
library(HypeRIMU)
data <- execute_TCP(port = 5555)
```

#### File

The recommended way to read the data is to save the records on your smartphone and submit them locally to your PC as a .csv file. The advantage from this method is the appropriation of sensor names. 

The following code shows an example of using a local file. The file can be imported via the function `execute_file()`. If your data provide a `timestamp`, choose `TRUE` for the argument `timestamp`. The functions tries to guess if a timestamp is available even when you set the `timestamp` argument wrong.

When a timestamp is available in the file, the function changes the UNIX format to as POSIXct format.


```{r}
library(HypeRIMU)
library(magrittr)

data <- system.file('extdata', 'short_y_impulse.csv', package = 'HypeRIMU') %>% 
  execute_file()
```

### Analyse sensor data

HypeRIMU provides a function to extract some specific sensor data, see next code.
When a sumbitted `sensorName` is not available in the `data`, an error message is printed.

```{r}
MPL_Accelerometer <- get_specificSensor(data, sensorName = "MPL_Accelerometer")

MPL_Linear_Acceleration <- get_specificSensor(data, sensorName = "MPL_Linear_Acceleration")
```

### Plot sensor data

Visulasing the data is one of the main advantages of this **R**-package. With some extra libraries a nice plot can be created in just a few lines. 
NOTE: It is recommended to use data with a timestamp.

```{r plot_sensor_data_1}
library(ggplot2)
library(reshape2)
library(scales)

melt(MPL_Linear_Acceleration, id.vars = "timestamp", variable.name = "Sensor") %>% 

ggplot(aes(x = timestamp, y = value, color = Sensor)) + 
  geom_line() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_x_datetime(labels = date_format("%H:%M:%S", tz = Sys.timezone()))
```


```{r plot_sensor_data2}
MPL_Accelerometer_melt <- melt(MPL_Accelerometer, id.vars = "timestamp", variable.name = "Sensor") %>% 

ggplot(aes(x = timestamp, y = value, color = Sensor)) + 
  geom_line() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_x_datetime(labels = date_format("%H:%M:%S", tz = Sys.timezone()))  
```


# Related projects 

* [HIMUServer](https://github.com/ianovir/HIMUServer/)
* [HyperIMU Android](https://play.google.com/store/apps/details?id=com.ianovir.hyper_imu)
* [Android API](https://developer.android.com/guide/topics/sensors/index.html)


# To Do

* support UDP stream
* make examples
* write function to get IP adress of server

# License

This program is free software: you can redistribute it and/or modify
it under the terms of the GNU General Public License as published by
the Free Software Foundation, either version 3 of the License, or
any later version.

This program is distributed in the hope that it will be useful,
but WITHOUT ANY WARRANTY; without even the implied warranty of
MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  
