---
title: "Getting started with HypeRIMU"
author: "Johannes Friedrich"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Getting started with HypeRIMU}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

# Motivation

Modern smartphones include a lot of sensors, e.g. for gravity or linear acceleration. Some apps provide the data collected from these sensors. One famous android app is "HyperIMU" (IMU: Inertial measurement unit). There are other apps available, but HyperIMU has some advantages:

* Choose which sensor data should be collected
* Choose different streaming protocols: SD-card (locally on the smarphone), UDP, or TCP
* The sampling rate can be adjusted

The **R**-package HypeRIMU provides some handy functions to visualise sensor data. Different streaming protocolls are supported (a local .csv file and TCP).

# Usage

For using the smartphone app, see [HIMUServer](https://github.com/ianovir/HIMUServer/) and [HyperIMU Android](https://play.google.com/store/apps/details?id=com.ianovir.hyper_imu&hl=de) and install the app on your smartphone.

## Read sensor data

Until now two different methods to read the sensor data are supported: TCP and reading a local file.

### TCP

The TCP protocol is also supported. Using this method needs to know your IP address to sumbit in the smartphone app, see [HIMUServer](https://github.com/ianovir/HIMUServer/).

```{r, eval=FALSE}
data <- execute_TCP(port = 5555, timestamp = T)
```

### File

The recommended way to read the data is to save the records on your smartphone and submit them locally to your PC as a .csv file. The advantage from this method is the appropriation of sensor names. 

The following code shows an example of using a local file. The file can be imported via the function `execute_file()`. If your data provide a `timestamp`, choose `TRUE` for the argument `timestamp`. The functions tries to guess if a timestamp is available even when you set the `timestamp` argument wrong.

When a timestamp is available in the file, the function changes the UNIX format to as POSIXct format.


```{r}
library(HypeRIMU)
file <- system.file('extdata', 'short_y_impulse.csv', package = 'HypeRIMU')
data <- execute_file(file = file, timestamp = TRUE)
```

## Analyse sensor data

HypeRIMU provides a function to extract some specific sensor data, see next code.
When a sumbitted `sensorName` is not available in the `data`, an error message is printed.

```{r}
MPL_Accelerometer <- get_specificSensor(data, sensorName = "MPL_Accelerometer")

MPL_Linear_Acceleration <- get_specificSensor(data, sensorName = "MPL_Linear_Acceleration")
```

## Plot sensor data

Visulasing the data is one of the main advantages of this **R**-package. With some extra libraries a nice plot can be created in just a few lines. 
NOTE: It is recommended to use data with a timestamp.

```{r plot sensor data}
library(ggplot2)
library(reshape2)
library(scales)

MPL_Linear_Acceleration_melt <- melt(MPL_Linear_Acceleration, id.vars = "timestamp")

ggplot(MPL_Linear_Acceleration_melt, aes(x = timestamp, y = value, color = variable)) + 
  geom_line() +
  scale_x_datetime(labels = date_format("%H:%M:%S")) +
  theme(legend.position = "bottom")

```

```{r}

MPL_Accelerometer_melt <- melt(MPL_Accelerometer, id.vars = "timestamp")

ggplot(MPL_Accelerometer_melt, aes(x = timestamp, y = value, color = variable)) + 
  geom_line() +
  scale_x_datetime(labels = date_format("%H:%M:%S"))+
  theme(legend.position = "bottom")
```
